---
title: "Reproducible Research: Peer Assessment 1"
output: 
  html_document:
    keep_md: true
---


## Loading and preprocessing the data
Unzipping the file, and then reading the data.

```{r}

unzip("repdata_data_activity.zip")
dataset <- read.csv("activity.csv")

```


## What is mean total number of steps taken per day?
2. Histogram - total steps taken each day:

```{r}

steps_date <- aggregate(steps ~ data, data = dataset, FUN = sum)
hist(steps_date$steps, names.arg = steps_date$date, xlab = "Date", ylab = "Steps Taken", main = "Total number of steps taken per day")

```
3. Mean and median of total number of steps:

```{r}

mean(steps_date$steps)
median(steps_date$steps)

```
## What is the average daily activity pattern?
1. Time series plot:
```{r}

interval <- aggregate(steps ~ interval, data = dataset, FUN = mean)
plot(interval, type="l", xlab = "Avg. number of steps", ylab = "Interval number", main = "Avg. number of steps per interval")

```

2. Which 5-minute interval, on average across all the days in the dataset, contains the maximum number of steps?
```{r}

interval$interval[which.max(interval$steps)]

```
Interval number 835 contained the max no. of average steps.
 
## Imputing missing values

1. Calculate and report the total number of missing values in the dataset (i.e. the total number of rows with **NA**)

```{r}

sum(is.na(dataset))

```

2. Devise a strategy for filling in all of the missing values in the dataset. The strategy does not need to be sophisticated.

Could replace the values with mean in the time interval.

3. Create a new dataset that is equal to the original dataset but with all the missing data filled in. 
```{r}

dataset <- merge(dataset, interval, by = "interval", suffixes = c("", ".aggregate"))

all-na <- is.na(dataset$steps)

dataset$steps[all-na] <- dataset$steps.aggregate[all-na]

dataset <- dataset[,c(1:3)]

```

4. Histogram of the totak number of steps taken each day:

```{r}

steps_date <- aggregate(steps ~ date, data = dataset, FUN = sum)
hist(steps_date$steps, names.arg = steps_date$date, xlab = "Date", ylab = "Steps", main = "Total number of steps taken each day")

```

The new mean is **mean** 
```{r}

mean(steps_date$steps)

```
and the new median is **median** 
```{r}

median(steps_date$steps)

```

The difference between the new mean, median and the old mean, median is not much. 

## Are there differences in activity patterns between weekdays and weekends?

1. Create a new factor variable in the dataset with two levels – “weekday” and “weekend” indicating whether a given date is a weekday or weekend day.

```{r, cache = TRUE}

day <- funcion(date){
  if(weekdays(as.Date(date)) %in% c("Saturday","Sunday")) {
    "Weekend"
  } else{
    "Weekday"
  }
}

dataset$day_type <- as.factor(sapply(dataset$date, day))
```

2. Making a panel plot containing a time series plot of the 5-minute interval and the average number of steps taken, averaged across all weekday days or weekend days.

```{r}

par(mfrow = c(2,1))

for (type in c("Weekend", "Weekday")) {
  steps_type <- aggregate(steps ~ interval, data = dataset, subset = (dataset$day_type==type), FUN = mean)
  plot(steps_type, type = "l", main = type)
}
```

